---
title: "TE processing"
author: "Chandler Sutherland"
date: "2023-01-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
library(openxlsx)
library(ggsignif)
library(ggpubr)
```

Goal: Determine the median TE distances between hv and non-hvNLRs, and if hvNLRs are more likely to overlap with TEs. 

```{r}
#load data, filter to NLRs
table <- readxl::read_xlsx(path="C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\all_gene_table.xlsx")

NLR_table <- table %>% filter(HV != 'all_genes') 

```

Calculate median 
```{r}
hv_med <- NLR_table[NLR_table$HV == 'hv', ][['te_dist']] %>% median()
nhv_med <- NLR_table[NLR_table$HV == 'non-hv', ][['te_dist']] %>% median()

paste('The median TE distance for hvNLRs is ', hv_med, 'while the median te distance for non-hvNLRs is', nhv_med)
```
What percent of hvNLRs overlap with a TE? And vice versa?
```{r}
hv_percent <- sum(NLR_table[NLR_table$HV == 'hv', ][['te_dist']] == 0)/length(NLR_table[NLR_table$HV == 'hv', ][['te_dist']])

nhv_percent <- sum(NLR_table[NLR_table$HV == 'non-hv', ][['te_dist']] == 0)/length(NLR_table[NLR_table$HV == 'non-hv', ][['te_dist']])

paste('The % of hvNLRs overlapping with a TE is ', hv_percent*100, 'while the % of non-hvNLRs overlapping with a TE is', nhv_percent*100)

```

Chi-square test of independence
Relationship betwen two categorical variables (overlapping or not). Null: variables are independent, knowing the value of one does not help predict the value of the other. 

H1: variables are dependent, there is a relationship between the categorical variables, knowing the value of one variable helps to predict the value of the other 


```{r}
#create a contingency table 
NLR_table$te_overlap <- ifelse(NLR_table$te_dist ==0, 'te overlap', 'not')
cont_tbl <- table(NLR_table$HV, NLR_table$te_overlap)

chisq.test(cont_tbl)
```


