---
title: "hv_plot"
author: "Chandler Sutherland"
date: "2023-09-11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
```

Copyright (c) Chandler Sutherland
Email: chandlersutherland@berkeley.edu

Purpose: generate a named entropy plot to demonstrate hv and non-hvNLRs

Intermediate processing steps are shown here, with all input files available for download from Zenodo. Figures can be recreated using just the numerical source data provided in `Source Data/Figure 1` and `/EV Figure 1`.

```{r}
library(tidyverse)
library(ggplot2)
library(stringr)
library(ggrepel)
```

```{r}
#load in Daniil's entropy by position table 
Atha_Ent <- read_delim("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\Zenodo V2\\Athaliana_NLR_Entropy.tsv")

#load in the gene table, which provides HV information by clade 
Gene_table <- read_delim("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\Zenodo V2\\Atha_NLRome_GeneTable.txt",delim = "\t")

#filter to just Col-0 
Col_Ent <- Atha_Ent %>% filter(str_detect(Name, 'ATHALIANA')) %>% separate(Name, c(NA, 'Gene'))
Col_table <- Gene_table %>% 
  filter(Ecotype=='ATHALIANA') %>% 
  separate(Gene, c(NA, 'Gene')) %>% 
  subset(select=c('Gene', 'Clade', 'HV')) %>% 
  distinct()

#read in common names 
names <- read_csv("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\Zenodo V2\\Atha_NLR_common_names.csv") 

#tack common names onto the subset gene table 
Col_meta <- Col_table %>% merge(names)

#add back on to the entropy table 
named_ent <- left_join(Col_Ent, Col_meta, by=c('Gene', 'Clade')) %>% drop_na(HV)
```
Fig 1: mean entropy per NLR 
```{r}
#create a df with the mean entropy
fig1_source <- named_ent %>% 
  group_by(Gene, HV, name) %>% 
  summarise(mean_ent = mean(Ent))

#save source data 
write.csv(fig1_source, file="./Source Data/Figure 1/Col0_mean_entropy.csv")
#read in source data 
#fig1_source <- read.csv("./Source Data/Figure 1/Col0_mean_entropy.csv")

#create a histogram
mean_entropy <- ggplot() + 
  geom_histogram(fig1_source, mapping=aes(x=mean_ent, fill=as.factor(-HV)), col=I('grey'))+
  xlab("Mean Entropy (bits)")+
  ylab("Number of NLRs")+
  theme_classic() +
  theme(text = element_text(size=10), 
        legend.position='none')+
  scale_fill_discrete(name ="", labels = c("hvNLR", "non-hvNLR"))

mean_entropy

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\Outputs\\NLR Features Paper\\EMBO Submission\\Figure Panels\\fig_1a.svg', plot=mean_entropy, dpi=1000, width=87, height=55, units='mm')
```

EV fig 1: count vs entropy at the tenth highest entropy position 
```{r}
#create a df with the tenth highest entropy position 
col_r10 <- named_ent %>% group_by(Gene) %>% slice_max(Ent,n=10) %>% slice_min(Ent,n=1,with_ties = FALSE) %>% 
  ungroup

#create a labelled histogram 
hist <- ggplot() + 
  geom_histogram(col_r10, mapping=aes(x=Ent, fill=as.factor(-HV)), col=I('grey'), bins=30)

#create a column, bins2, that allows for plotting of label and create label_df 
real_bin <- ggplot_build(hist)$data[[1]] %>% subset(select=c(y, xmin, xmax))
breaks <- real_bin$xmin %>% unique()
plot_df <- col_r10 %>% mutate(bins2=as.factor(as.numeric(cut(Ent, breaks=breaks))))%>% arrange(Ent)

#save source data 
ev_fig1_source <- plot_df %>% subset(select=c('Gene', 'HV', 'name', 'Ent', 'bins2'))
write.csv(ev_fig1_source, file='./Source Data/Figure 1/Col0_mean_entropy.csv')
#ev_fig1_source <- read.csv(file='./Source Data/Figure 1/Col0_mean_entropy.csv')

#create label_df
counts <- ev_fig1_source %>% group_by(bins2) %>% count
label_df <- ev_fig1_source %>% filter(!is.na(name)) %>% left_join(counts, by='bins2')

#plot 
r10 <- ggplot() + 
  geom_histogram(ev_fig1_source, mapping=aes(x=Ent, fill=as.factor(-HV)), col=I('grey'), bins=30)+
  geom_vline(xintercept = 1.5,linetype="dotted")+
  geom_label_repel(label_df, 
                   mapping=aes(label=name, x=Ent, y=n, color=as.factor(-HV)), 
                   nudge_y=5, show.legend = FALSE, segment.color='grey')+
  xlab("Entropy (bits) at the 10th highest position")+
  ylab("Number of NLRs")+
  theme_classic() +
  theme(legend.position = 'none')+
  scale_fill_discrete(name ="", labels = c("HV", "non-HV"))

r10

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\Outputs\\NLR Features Paper\\EMBO Submission\\Figure Panels\\EVfig_1.svg', plot=r10, dpi=1000, width=180, height=87, units='mm')
```



