---
title: "R Notebook"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Copyright (c) Chandler Sutherland
Email: chandlersutherland@berkeley.edu

Purpose: generate iTOL annotation files corresponding to the tree 
```{r}
library(tidyverse)
library(ggplot2)
library(openxlsx)
library(ggsignif)
library(ggpubr)
```

First, add the itol branch length names and common gene names to the methylation/expression/TE distance information 
```{r}
#read in the methylation, expression, and TE distance for all genes, filter to NLRs
zenodo_path <- "C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\Zenodo V2\\"
NLR_table <- read.csv(paste(zenodo_path, 'NLR_gene_table.csv'))

#add gene name information and the itol branch labels 
gene_names <- readxl::read_xlsx(path = "C://Users//chand//Box Sync//Krasileva_Lab//Research//chandler//Krasileva Lab//E14//gene_names_Daniil_tree.xlsx", col_names=c('itol_name', 'gene_name'))

names <- gene_names %>% separate(col=itol_name, sep=' ', into=c(NA, 'transcript_name'), remove=FALSE) %>% filter(itol_name != 'I264') %>% filter(itol_name != 'I0')
names$Gene <- substr(names$transcript_name, 1, nchar(names$transcript_name)-2)

named_table <- merge(names, NLR_table, by='Gene', all.x=TRUE)

```

Starting with some binaries: HV/nonhv
```{r}
itol_HV <- named_table[c('itol_name', 'HV')]
itol_HV <- itol_HV %>% mutate(HV_itol=recode(HV, 'hv'=1, 'non-hv'=-1)) %>% subset(select=c('itol_name', 'HV_itol'))
write.xlsx(itol_HV, 'C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\nlr_features\\figure_generation\\Source Data\\Figure 3\\3B\\itol_HV.xlsx')
```

Color gradient for expression and methylation
```{r}
itol_TPM <- named_table[c('itol_name', 'log2_TPM')]
write.xlsx(itol_TPM, 'C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\nlr_features\\figure_generation\\Source Data\\Figure 3\\3B\\itol_tpm.xlsx')

itol_meth <- named_table[c('itol_name', 'meth_percentage')]
write.xlsx(itol_meth, 'C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\nlr_features\\figure_generation\\Source Data\\Figure 3\\3B\\itol_meth.xlsx')
```


TE could easily become a binary, which may be more informative anyway 
```{r}
#how about TE within 1kb
named_table['te_neighbor_500'] = named_table['te_dist'] <= .5
named_table$te_neighbor_500 <- named_table$te_neighbor_500 %>% as.integer() %>% as.character()

itol_te_500 <- named_table[c('itol_name', 'te_neighbor_500')]
itol_te_500 <- itol_te_500 %>% mutate(te_neighbor_500=recode(te_neighbor_500, '0'='-1'))
itol_te_500[is.na(itol_te_500)] <- '1'

write.xlsx(itol_te_500, 'C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\nlr_features\\figure_generation\\Source Data\\Figure 3\\3B\\itol_te_500.xlsx')
```


