---
title: "R Notebook"
output: html_notebook
---
Author: Chandler Sutherland 
Copyright (c) Chandler Sutherland
Email: chandlersutherland@berkeley.edu

Goal: assign significance to the enrichment plots, using the singscore gene set enrichment for expression, and permutation tests for methylation and TE distance. Also, use permutations to determine number of NLRs expected by chance in tails of empirical distribution. 

```{r}
library(data.table)
library(ggplot2)
library(tidyverse)
library(openxlsx)
library(ggsignif)
library(ggpubr)
library(singscore)
library(GSEABase)
```

Read in the gene table 
```{r}
table <- readxl::read_xlsx(path="C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\all_gene_table.xlsx")
grey <- read.csv("C://Users//chand//Box Sync//Krasileva_Lab//Research//chandler//Krasileva Lab//E8//Ath_somatic_mutations_gene_level_data.csv")
merged <- merge(grey, table  %>% subset(select=c('Gene', 'HV')), by.x='gene', by.y='Gene', all=TRUE) %>% 
  mutate(HV=recode(HV, `0` = "non-hv", `1`="hv")) %>% rename('Gene'='gene')
```

An out of the box package exists to test significance of gene sets for expression: singscore. (Fourtran 2018)
```{r}
#generate gene sets
#remove unmappable genes 
HV <- table %>% filter(HV == 'hv') %>% filter(!(Gene %in% c('AT1G58807','AT1G58848','AT1G59124','AT1G59218')))
hv_names <- HV$Gene
hv_geneset <- GeneSet(hv_names, geneIdType=SymbolIdentifier())

NONHV <- table %>% filter(HV == 'non-hv')
nonhv_names <- NONHV$Gene %>% unique()
nonhv_geneset <- GeneSet(nonhv_names, geneIdType=SymbolIdentifier())
```

This packages does analysis by sample, not on averaged biological replicates. Loading a raw TPM matrix. Then, create the ranked gene object, ranked, and score the hv gene set and the nonhv gene set. From here, generate a null distro and plot our p values 
```{r}
TPM <- read.table("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\tpm_matrix.csv", header=TRUE, sep=",", row.names=2) %>% subset(select=c(-X))

ranked <- rankGenes(TPM)

hv_sing <- simpleScore(ranked, upSet=hv_geneset)
nonhv_sing <- simpleScore(ranked, nonhv_geneset)

hv_null <- generateNull(rankData=ranked, upSet=hv_geneset)
hv_p <- getPvals(hv_null, hv_sing)
plotNull(hv_null, hv_sing, hv_p, sampleNames = names(hv_p)[1:4], textSize=0.5)
```

```{r}
nonhv_null <- generateNull(rankData=ranked, upSet=nonhv_geneset)
nonhv_p <- getPvals(nonhv_null, nonhv_sing)
plotNull(nonhv_null, nonhv_sing, nonhv_p, sampleNames = names(nonhv_p)[1:4], textSize=0.5)
```

For methylation and TE distance, write a function to calculate a p value for enrichment tests by using a permutation test 
```{r}
# first, calculate the observed value
observe <- function(df, col_index, hv_status){
  observed <- mean(df%>%filter(HV==hv_status)%>% pull(col_index)%>%na.omit()) - 
    mean(df %>% filter(HV=='all_genes')%>%pull(col_index)%>%na.omit())
  observed
}

permute <- function(df, col_index, hv_status) {
  # build null distribution
  permutation = replicate(10000, {
    df <- df %>% drop_na(col_index)
    sample_small <- df[sample(nrow(df), nrow(df[df$HV == hv_status,]) , replace = FALSE), ] # sample the size of hvNLRs
    sample_large <- df[! rownames(df) %in% rownames(sample_small), ] # sample the rest
    mean(sample_small %>% pull(col_index))-mean(sample_large %>% pull(col_index)) # get expected mean
})
  permutation
}


p_calc <- function(df, col_index){
  o <- observe(df, col_index, 'hv')
  p <- permute(df, col_index, 'hv')
  pval <- mean(o > p)
  
  o2 <- observe(df, col_index, 'non-hv')
  p2 <- permute(df, col_index, 'non-hv')
  pval2 <- mean(o2 > p2)
  
  print(paste(colnames(df)[col_index]))
  print(paste('hv_p: ', pval, ' nonhv_p:', pval2))
}

p_calc(table, 4)
p_calc(table, 5)
p_calc(table, 6)
p_calc(table, 7)
p_calc(table, 8)
p_calc(merged, 35)
```
```{r}
permute_hv <- permute(table, 4, 'hv')
q=quantile(permute_hv, c(.95, .05))
observe_hv <- observe(table, 4, 'hv')
ggplot() + aes(permute_hv)+ geom_histogram(colour="black", fill="white") + geom_vline(xintercept=observe_hv)+
  geom_vline(aes(xintercept=q[2]), colour="#BB0000", size=1)+
  geom_vline(aes(xintercept=q[1]), colour="#BB0000", size=1)+
  #geom_vline(aes(xintercept=observed_nhv), color='blue')+
  xlab('Permutation difference in means meth hv nlr 1000')

mean(observe_hv > permute_hv)
```

```{r}
table <- table %>% drop_na(D)
observed_nhv <- mean(table$D[table$HV == 'non-hv' ]) - mean(table$D[table$HV == 'all_genes' ])

permutation_nhv = replicate(1000, {
  sample_small <- table[sample(nrow(table), nrow(table[table$HV == 'non-hv',]) , replace = FALSE), ] # sample the size of non-hvNLRs
  sample_large <- table[! rownames(table) %in% rownames(sample_small), ] # sample the rest
  mean(sample_small$D)-mean(sample_large$D) # get expected mean
})
q=quantile(permutation_nhv, c(.05, .95))
#did this by hand, but want to plot p=.05 on the histo 

p <- ggplot() + aes(permutation_nhv)+ geom_histogram(colour="black", fill="white") + geom_vline(xintercept=observed_nhv)+
  geom_vline(aes(xintercept=q[2]), colour="#BB0000", size=1)+
  geom_vline(aes(xintercept=q[1]), colour="#BB0000", size=1)+
  geom_vline(aes(xintercept=observed_nhv), color='blue')+
  xlab('Permutation D non-hv 1000')
p

observed_hv <- mean(table$D[table$HV == 'hv' ]) - mean(table$D[table$HV == 'all_genes' ])

permutation_hv = replicate(1000, {
  sample_small <- table[sample(nrow(table), nrow(table[table$HV == 'hv',]) , replace = FALSE), ] # sample the size of non-hvNLRs
  sample_large <- table[! rownames(table) %in% rownames(sample_small), ] # sample the rest
  mean(sample_small$D)-mean(sample_large$D) # get expected mean
})
q=quantile(permutation_hv, c(.05, .95))
#did this by hand, but want to plot p=.05 on the histo 

p <- ggplot() + aes(permutation_hv)+ geom_histogram(colour="black", fill="white") + geom_vline(xintercept=observed_hv)+
  geom_vline(aes(xintercept=q[2]), colour="#BB0000", size=1)+
  geom_vline(aes(xintercept=q[1]), colour="#BB0000", size=1)+
  #geom_vline(aes(xintercept=observed_nhv), color='blue')+
  xlab('Permutation D hv 1000')
p

#check all NLRs vs Pi 
observed_nlr <- mean(table$Pi_by_lseff[table$HV != 'all_genes' ]) - mean(table$Pi_by_lseff[table$HV == 'all_genes' ], na.rm=TRUE)

permutation_nlr = replicate(1000, {
  sample_small <- table[sample(nrow(table), nrow(table[table$HV != 'all_genes',]) , replace = FALSE), ] # sample the size of NLRs
  sample_large <- table[! rownames(table) %in% rownames(sample_small), ] # sample the rest
  mean(sample_small$Pi_by_lseff)-mean(sample_large$Pi_by_lseff) # get expected mean
})
q=quantile(permutation_nlr, c(.05, .95))
#did this by hand, but want to plot p=.05 on the histo 

p <- ggplot() + aes(permutation_nlr)+ geom_histogram(colour="black", fill="white") + geom_vline(xintercept=observed_nlr)+
  geom_vline(aes(xintercept=q[2]), colour="#BB0000", size=1)+
  geom_vline(aes(xintercept=q[1]), colour="#BB0000", size=1)+
  #geom_vline(aes(xintercept=observed_nhv), color='blue')+
  xlab('Permutation Pi NLR 1000')
p

mean(observed_nlr < permutation_nlr)
```


For TE distance, also check the median: 
```{r}
observe_med <- function(df, col_index, hv_status){
  observed <- median(df%>%filter(HV==hv_status)%>%pull(col_index)%>%na.omit()) - 
    median(df %>% filter(HV=='all_genes')%>%pull(col_index)%>%na.omit())
  observed
}

permute_med <- function(df, col_index, hv_status) {
  # build null distribution
  permutation = replicate(10000, {
    df <- df %>% drop_na(col_index)
    sample_small <- df[sample(nrow(df), nrow(df[df$HV == hv_status,]) , replace = FALSE), ] # sample the size of hvNLRs
    sample_large <- df[! rownames(df) %in% rownames(sample_small), ] # sample the rest
    median(sample_small %>% pull(col_index))-median(sample_large %>% pull(col_index)) # get expected mean
})
  permutation
}

p_calc_med <- function(df, col_index){
  o <- observe_med(df, col_index, 'hv')
  p <- permute_med(df, col_index, 'hv')
  pval <- mean(o > p)
  
  o2 <- observe_med(df, col_index, 'non-hv')
  p2 <- permute_med(df, col_index, 'non-hv')
  pval2 <- mean(o2 > p2)
  
  print(paste(colnames(df)[col_index], ' median'))
  print(paste('hv_p: ', pval, ' nonhv_p:', pval2))
}

p_calc_med(table, 5)
```


A potential bias for CG methylation is the number of high coverage CGs per gene. This is on a per-sample basis. First, test the per sample significance, then weight based on CGs per gene. 
```{r}
#load the per sample methylation files 
SRR17281085 <- readxl::read_xlsx(path="C://Users//chand//Box Sync//Krasileva_Lab//Research//chandler//Krasileva Lab//E14//e14_R//SRR17281085_per_gene_meth_count.xlsx", skip=2, col_names = c('Index', 'Chrom', 'Gene', 'mean_percent_methylation', 'count'))
SRR17281086 <- readxl::read_xlsx(path="C://Users//chand//Box Sync//Krasileva_Lab//Research//chandler//Krasileva Lab//E14//e14_R//SRR17281086_per_gene_meth_count.xlsx", skip=2, col_names = c('Index', 'Chrom', 'Gene', 'mean_percent_methylation', 'count'))
SRR17281087 <- readxl::read_xlsx(path="C://Users//chand//Box Sync//Krasileva_Lab//Research//chandler//Krasileva Lab//E14//e14_R//SRR17281087_per_gene_meth_count.xlsx", skip=2, col_names = c('Index', 'Chrom', 'Gene', 'mean_percent_methylation', 'count'))
SRR17281088 <- readxl::read_xlsx(path="C://Users//chand//Box Sync//Krasileva_Lab//Research//chandler//Krasileva Lab//E14//e14_R//SRR17281088_per_gene_meth_count.xlsx", skip=2, col_names = c('Index', 'Chrom', 'Gene', 'mean_percent_methylation', 'count'))

#clean up 
samples <- list(SRR17281085, SRR17281086, SRR17281087, SRR17281088)
names(samples) <- c('SRR17281085', 'SRR17281086', 'SRR17281087', 'SRR17281088')

clean <- function(df) {
  df2 <- df %>% merge(table, by='Gene', all=TRUE)
  df2$HV <- factor(df2$HV , levels=c("all_genes", "non-hv", "hv"))
  df3 <- df2 %>% subset(select=c('Gene', 'HV', 'mean_percent_methylation', 'count')) %>%
    filter(!is.na(mean_percent_methylation))
  df3
}

clean_samples <- lapply(names(samples), function(x) clean(samples[[x]]))
names(clean_samples) <- c('SRR17281085', 'SRR17281086', 'SRR17281087', 'SRR17281088')
```
```{r}
#lapply(names(clean_samples), function(x) p_calc(clean_samples[[x]], 3))
```

The kernel density estimation computes kernel density estimates. Often shortened to KDE, it’s a technique that let’s you create a smooth curve given a set of data. So here, it's taking in my # of mappable CG sites per HV gene, and returning a density curve. 

The approxfun function returns a list of points which linearly interpolate given data points. Returns a function performing linear or constant interpolation of the given data points. Linear interpolation is a method of calculating intermediate data between known values by conceptually drawing a straight line between two adjacent known values.

Repeat the CG count weighted permutation test on all four samples: 
```{r}
#create a sampled gene list weighted by counts 
density_sample <- function(df, hv_status){
  #df <- df %>% drop_na()
  split_df <- df %>% filter(HV == hv_status)
  
  #calculate CG count density KDE function 
  d <- density(split_df$count)
  density_function <- approxfun(d)
  
  # write probabilities, and write NAs to zero as they are outliers in CG content (90+ per gene)
  probabilities <- density_function(df$count)
  probabilities[is.na(probabilities)] <- 0
  
  # sample genes, with weights according to their number of CGs, according to above function
  all_genes <- df$Gene
  sampled_genes <- data.frame(replicate(10000, sample(all_genes, size = nrow(split_df), replace = FALSE, 
                                                    prob = probabilities)))
  
  return(sampled_genes)
  }

test_nhv <- density_sample(clean_samples[[1]], 'non-hv')
table <- clean_samples[[1]]

#write function that gets the difference in means between the permuted sample and the actual 
get_mean_difference <- function(gene_list, table) {
  permute_mean <- table %>% filter(Gene %in% gene_list) %>% pull(mean_percent_methylation) %>% mean()
  actual_mean <- table %>% filter(!Gene %in% gene_list)  %>% pull(mean_percent_methylation) %>% mean()
  return(actual_mean-permute_mean)
}


get_mean_difference(test_nhv[[1]], clean_samples[[1]])
mean_diff <- actual_mean-permute_mean

#calculate the pvalue 
pval <- function(sampled_df, table, hv_status, observed){
  differences <- list()
  for (x in 1:10000) {
    x <- get_mean_difference(sampled_df[[x]], table)
    differences <- append(differences, x)
    }
  
  p_val <- mean(differences < observed)
  
  return(p_val)
}

pval(test_nhv, 'non-hv', observed_nhv)
```

```{r}
hv_sampled_genes <- lapply(names(clean_samples), function(x) density_sample(clean_samples[[x]], 'hv'))
observed_hv <- lapply(names(clean_samples), function(x) observe(clean_samples[[x]], 3, 'hv'))

pval(hv_sampled_genes[[1]], clean_samples[[1]], 'hv', observed_hv[[1]])
pval(hv_sampled_genes[[2]], clean_samples[[2]], 'hv', observed_hv[[2]])
pval(hv_sampled_genes[[3]], clean_samples[[3]], 'hv', observed_hv[[3]])
pval(hv_sampled_genes[[4]], clean_samples[[4]], 'hv', observed_hv[[4]])
```

```{r}
nhv_sampled_genes <- lapply(names(clean_samples), function(x) density_sample(clean_samples[[x]], 'non-hv'))
observed_nhv <- lapply(names(clean_samples), function(x) observe(clean_samples[[x]], 3, 'non-hv'))

pval(nhv_sampled_genes[[1]], clean_samples[[1]], 'non-hv', observed_nhv[[1]])
pval(nhv_sampled_genes[[2]], clean_samples[[2]], 'non-hv', observed_nhv[[2]])
pval(nhv_sampled_genes[[3]], clean_samples[[3]], 'non-hv', observed_nhv[[3]])
pval(nhv_sampled_genes[[4]], clean_samples[[4]], 'non-hv', observed_nhv[[4]])
```

```{r}
#function that takes in table, column name as a string, and n replicates, and outputs probability of observing NLRs, hvNLRs, and non-hvNLRs in the top 5% of the distribution by chance 
on_top <- function(table, x, n){
  
  q <- quantile(table %>% pull(x), c(.05, .95), na.rm=TRUE)
  print(paste('calculating probability of ', x, '. The 95th percentile is: ', q[2]))
  
  observed_nlr <- table[table['HV'] != 'all_genes' & table[x]>q[2],] %>% nrow()
  
  permutation_nlr = replicate(n, {
  sample_small <- table[sample(nrow(table), nrow(table[table$HV != 'all_genes',]) , replace = FALSE), ] # sample the size of all NLRs
  sample_small[sample_small[x]>q[2],] %>% nrow()  # get expected above 95% 
  })
  
  print(paste('observed NLRs in top 5%:', observed_nlr))
  print(paste('expected NLRs in top 5% by chance:', mean(permutation_nlr)))
  print(paste('p value non-hv in top 5%:', mean(observed_nlr < permutation_nlr)))
  
  observed_hv <- table[table['HV'] == 'hv' & table[x]>q[2],] %>% nrow()
  
  permutation_hv = replicate(n, {
  sample_small <- table[sample(nrow(table), nrow(table[table$HV == 'hv',]) , replace = FALSE), ] # sample the size of hvNLRs
  sample_small[sample_small[x]>q[2],] %>% nrow()  # get expected above 95% 
  })
  
  print(paste('observed hvNLRs in top 5%:', observed_hv))
  print(paste('expected hvNLRs in top 5% by chance:', mean(permutation_hv)))
  print(paste('p value hv in top 5%:', mean(observed_hv < permutation_hv)))
  
  observed_nhv <- table[table['HV'] == 'non-hv' & table[x]>q[2],] %>% nrow()
  
  permutation_nhv = replicate(n, {
  sample_small <- table[sample(nrow(table), nrow(table[table$HV == 'non-hv',]) , replace = FALSE), ] # sample the size of non-hvNLRs
  sample_small[sample_small[x]>q[2],] %>% nrow()  # get expected above 95% 
  })
  
  print(paste('observed hvNLRs in top 5%:', observed_nhv))
  print(paste('expected hvNLRs in top 5% by chance:', mean(permutation_nhv)))
  print(paste('p value hv in top 5%:', mean(observed_hv < permutation_nhv)))
  
}

on_top(table, 'Pi_by_lseff', 10000)
on_top(table, 'D', 10000)

```
```{r}
on_bottom <- function(table, x, n){
  
  q <- quantile(table %>% pull(x), c(.05, .95), na.rm=TRUE)
  print(paste('calculating probability of ', x, '. The 5th percentile is: ', q[1]))
  
  observed_nlr <- table[table['HV'] != 'all_genes' & table[x]<q[1],] %>% nrow()
  
  permutation_nlr = replicate(n, {
  sample_small <- table[sample(nrow(table), nrow(table[table$HV != 'all_genes',]) , replace = FALSE), ] # sample the size of all NLRs
  sample_small[sample_small[x]<q[1],] %>% nrow()  # get below 5% 
  })
  
  print(paste('observed NLRs in bottom 5%:', observed_nlr))
  print(paste('expected NLRs in bottom 5% by chance:', mean(permutation_nlr)))
  print(paste('p value non-hv in bottom 5%:', mean(observed_nlr < permutation_nlr)))
  
  observed_hv <- table[table['HV'] == 'hv' & table[x]<q[1],] %>% nrow()
  
  permutation_hv = replicate(n, {
  sample_small <- table[sample(nrow(table), nrow(table[table$HV == 'hv',]) , replace = FALSE), ] # sample the size of hvNLRs
  sample_small[sample_small[x]<q[1],] %>% nrow()  # get expected above 95% 
  })
  
  print(paste('observed hvNLRs in bottom 5%:', observed_hv))
  print(paste('expected hvNLRs in bottom 5% by chance:', mean(permutation_hv)))
  print(paste('p value hv in bottom 5%:', mean(observed_hv < permutation_hv)))
  
  observed_nhv <- table[table['HV'] == 'non-hv' & table[x]<q[1],] %>% nrow()
  
  permutation_nhv = replicate(n, {
  sample_small <- table[sample(nrow(table), nrow(table[table$HV == 'non-hv',]) , replace = FALSE), ] # sample the size of non-hvNLRs
  sample_small[sample_small[x]<q[1],] %>% nrow()  # get expected above 95% 
  })
  
  print(paste('observed hvNLRs in bottom 5%:', observed_nhv))
  print(paste('expected hvNLRs in bottom 5% by chance:', mean(permutation_nhv)))
  print(paste('p value hv in bottom 5%:', mean(observed_hv < permutation_nhv)))
  
}

on_bottom(table, 'Pi_by_lseff', 1000)
on_bottom(table, 'D', 1000)
```

Supplemental figure 3, which shows the distribution of D and Pi with the percentiles
```{r}
d_distro <- ggplot(data=table)+geom_histogram(aes(x=D))+
   geom_segment(x=1.50913785453558, xend=1.50913785453558, yend=0, y=20000, aes(color='5th and 95th percentiles'))+
  geom_segment(x=-1.7870429615165, xend=-1.7870429615165, yend=0, y=20000, aes(color='5th and 95th percentiles'))+
  geom_segment(data= table %>% filter(HV=='hv'), aes(
      x= D,
  		y = 0,
  		xend = D,
  		yend = -250, color='hvNLR'), alpha = 0.75, linewidth=0.5) +
  	labs(colour = c('hvNLR', 'nonhvNLR')) +
geom_segment(data= table %>% filter(HV=='non-hv'), aes(
      x= D,
  		y = -250,
  		xend = D,
  		yend = -500, color='non-hvNLR'), alpha = 0.75, linewidth=0.5) +
  	labs(colour = c('hvNLR', 'nonhvNLR')) +
  theme_classic() +
  theme(legend.position='right', text=element_text(size=10), legend.title=element_blank())+
  labs(title='Empirical Distribution of D values')+
  xlab("Tajima's D")+
  ylab('count')+
  scale_color_manual(values=c('red', '#F8766D', '#00BFC4'))

pi_distro <- ggplot(data=table)+geom_histogram(aes(x=Pi_by_lseff), bins=50)+
  geom_segment(x=0.000316624887148824, xend=0.000316624887148824, yend=0, y=20000, aes(color='5th and 95th percentiles'))+
  geom_segment(x=0.0140259207477425, xend=0.0140259207477425, yend=0, y=20000, aes(color='5th and 95th percentiles'))+
  geom_segment(data= table %>% filter(HV=='hv'), aes(
      x= Pi_by_lseff,
  		y = 0,
  		xend = Pi_by_lseff,
  		yend = -500, color='hvNLR'), alpha = 0.75, linewidth=0.5) +
  	labs(colour = c('hvNLR', 'nonhvNLR')) +
geom_segment(data= table %>% filter(HV=='non-hv'), aes(
      x= Pi_by_lseff,
  		y = -500,
  		xend = Pi_by_lseff,
  		yend = -1000, color='non-hvNLR'), alpha = 0.75, linewidth=0.5) +
  	labs(colour = c('hvNLR', 'nonhvNLR')) +
  theme_classic() +
  xlim(0, 0.10)+
  theme(legend.position='right', text=element_text(size=10), legend.title=element_blank())+
  labs(title='Empirical Distribution of Pi Values')+
  xlab(expression(pi))+
  ylab('count')+
  scale_color_manual(values=c('red', '#F8766D', '#00BFC4'))

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\d_distro.png', plot=d_distro, dpi='retina', width=5, height=3)

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\pi_distro.png', plot=pi_distro, dpi='retina', width=5, height=3)
```

Repeat for the density of both distributions 
```{r}
#double double toil and trouble, two distros! 
q_D <- quantile(table$D, c(.05, .95), na.rm = TRUE)
q_Pi <- quantile(table$Pi_by_lseff, c(.05, .95), na.rm=TRUE)

observed_nlr <- table %>% filter(HV != 'all_genes') %>% filter(D > q_D[2]) %>% filter(Pi_by_lseff > q_Pi[2]) %>% nrow()

permutation_nlr = replicate(10000, {
  sample_small <- table[sample(nrow(table), nrow(table[table$HV != 'all_genes',]) , replace = FALSE), ] # sample the size of non-hvNLRs
  #sample_large <- meth[! rownames(meth) %in% rownames(sample_small), ] # sample the rest
  sample_small%>% filter(D > q_D[2]) %>% filter(Pi_by_lseff > q_Pi[2]) %>% nrow()  # get expected above 95%
})

print(paste('observed NLRs in top 5% of both distros:', observed_nlr))
print(paste('expected NLRs in top 5% of both distros by chance:', mean(permutation_nlr)))
print(paste('p value NLRs in top of both distros:', mean(observed_nlr < permutation_nlr)))
```


