---
title: "popgen_plots"
author: "Chandler Sutherland"
date: "2023-08-15"
output: html_document
---

Purpose of script: Plot population genetics statistics of hv and non-hvNLRs  

Copyright (c) Chandler Sutherland
Email: chandlersutherland@berkeley.edu

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(ggplot2)
library(ggsignif)
library(ggpubr)
library(ggbeeswarm)
require(tidyverse)
library(openxlsx)
library(patchwork)
```

Gather and clean the necessary datasets 
```{r}
#read in col0_popgen_stats, generated by the egglib_nlr.ipynb script 
col_popgen <- read_tsv("//wsl.localhost//Ubuntu//home//chandlersutherland//scratch//egglib_results.tsv") %>%
  mutate(Pi_bp=Pi_raw/dna_aln)%>%
  mutate(Pi_site=Pi_raw/lseff)%>%
  mutate(PiNPiS = PiN/PiS) %>%
  mutate(domain=recode(domain, 'cc' = "CC", 'tir'="TIR", 'nbarc'='NBARC', 'lrr'='LRR', 'cds'='CDS'))

col_popgen$domain <- factor(col_popgen$domain, levels=c('CDS', 'CC', 'TIR', 'NBARC', 'LRR'))
col_popgen$HV <- factor(col_popgen$HV, levels=c('non-hv', 'hv'))

#add in gene names 
names <- read_tsv("//wsl.localhost//Ubuntu//home//chandlersutherland//scratch//col0_per_domain_stats.tsv") %>% 
  dplyr::select(c('Gene', 'name')) %>% 
  distinct()
col_popgen <- merge(col_popgen, names, all.x=TRUE)

popgen_per_domain <- col_popgen %>% subset(select=c('Gene', 'D', 'PiN', 'PiS', 'clade', 'dna_aln', 'domain', 'lseff', 'HV', 'Pi_site', 'PiNPiS', 'name'))
write.xlsx(popgen_per_domain, 'C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\popgen_per_domain.xlsx')
```


```{r}
#filter to just CDS and add sample sizes to the label 
to_plot <- col_popgen %>% 
  filter(domain=='CDS')%>%
  group_by(HV) %>%
  mutate(n=n()) %>%
  mutate(label = paste0(HV,"\n", "n=",n))

labels <- to_plot$label %>% unique()
to_plot$label <- factor(to_plot$label, levels=c(labels[1], labels[2]))
```


```{r}
#read in sliding window analysis 
x <- to_plot %>% subset(select=c('Gene', 'dna_aln')) %>% dplyr::rename('gene_length'='dna_aln')
vig_slider <- read_csv("//wsl.localhost//Ubuntu//home//chandlersutherland//scratch//egglib_window_stats.csv") %>% 
  mutate(Pi_site=Pi_raw/lseff) %>% 
  mutate(nuc_midpoint = (start*3+dna_aln/2)) 

col_slider <- merge(x, vig_slider, by='Gene', all.x=FALSE) %>% mutate(normal=nuc_midpoint/gene_length) 

#read in Grey's mutation data 
grey <- read.csv("C://Users//chand//Box Sync//Krasileva_Lab//Research//chandler//Krasileva Lab//E8//Ath_somatic_mutations_gene_level_data.csv")
table <- readxl::read_xlsx(path="C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\all_gene_table.xlsx")
merged <- merge(grey, table  %>% subset(select=c('Gene', 'HV')), by.x='gene', by.y='Gene', all=TRUE) %>% 
  mutate(HV=recode(HV, `0` = "non-hv", `1`="hv"))

merged$HV[is.na(merged$HV)]<-'all_genes'
merged$HV <- factor(merged$HV , levels=c("all_genes", "non-hv", "hv"))
NLR_only <- merged %>% filter(HV != 'all_genes')

NLR_only %>% subset(select=c('gene', 'Mutation.Probability.Score')) %>% merge(to_plot, by.x='gene', by.y='Gene') %>% 
  write.csv("C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\popgen_NLR.csv")

#read in all gene data 
all_gene <- readxl::read_xlsx(path="C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\all_gene_table.xlsx")


```


```{r}
#read in hyphy positive selection results 
fel_results<-read_csv("//wsl.localhost//Ubuntu//home//chandlersutherland//scratch//fel_results.csv")%>%
  dplyr::rename('Clade'='clade')

meme_results<-read_csv("//wsl.localhost//Ubuntu//home//chandlersutherland//scratch//meme_results.csv")%>%
  dplyr::rename('Clade'='clade')

clades <- read.table(file='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\Atha_NLRome_GeneTable.txt', header=1) %>% 
  filter(Ecotype=='ATHALIANA') %>% 
  separate(Gene, c(NA, 'Gene'))%>%
  subset(select=c('Gene', 'Clade', 'HV')) %>% 
  distinct() %>% 
  mutate(HV=recode(HV, '0'='non-hv', '1'='hv'))

pos_sites <- merge(clades, fel_results %>% filter(direction == 'Pos.'), all.x=TRUE) %>% 
  merge(meme_results, by='Clade', all.x=TRUE) %>% 
  subset(select=c('Clade', 'Gene', 'HV', 'prop_fel_int_95', 'prop_meme_internal_95'))

write.xlsx(pos_sites, 'C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\positive_selection.xlsx')

pos_sites$HV <- factor(pos_sites$HV , levels=c("non-hv", "hv"))

sites_internal_long <- pos_sites %>% subset(select=c('Gene', 'HV', 'prop_fel_int_95', 'prop_meme_internal_95')) %>% pivot_longer(
    cols = c(prop_fel_int_95, prop_meme_internal_95), 
    names_to = "model",
    values_to = "value") %>%
  mutate(model=recode(model, 
                      'prop_fel_int_95' = "Pervasive", 
                      'prop_meme_internal_95' = 'Episodic'))
```


# Figure 3: 
Per domain plots 
Pi stats: 
```{r}
cds <- compare_means(Pi_site~HV, col_popgen %>% filter(domain=='CDS'), method='wilcox.test', paired=FALSE)$p
cc <- compare_means(Pi_site~HV, col_popgen %>% filter(domain=='CC'), method='wilcox.test', paired=FALSE)$p
tir <- compare_means(Pi_site~HV, col_popgen %>% filter(domain=='TIR'), method='wilcox.test', paired=FALSE)$p
nbarc <- compare_means(Pi_site~HV, col_popgen %>% filter(domain=='NBARC'), method='wilcox.test', paired=FALSE)$p
lrr <- compare_means(Pi_site~HV, col_popgen %>% filter(domain=='LRR'), method='wilcox.test', paired=FALSE)$p

pvalues <- c(cds, cc, tir, nbarc, lrr)
pvalues_adjust <- p.adjust(pvalues, method = 'BH', n = length(pvalues))
print(pvalues_adjust)

compare_means(Pi_site~domain, col_popgen %>% filter(HV=='hv'), method='wilcox.test', paired=FALSE)
compare_means(Pi_site~domain, col_popgen %>% filter(HV=='non-hv'), method='wilcox.test', paired=FALSE)
```
```{r}
pi_exp <- expression(pi)
p7 <- ggplot(col_popgen)+
  geom_boxplot(aes(x=domain, y=Pi_site, fill=HV), lwd=0.25, outlier.size=0.1)+
    scale_fill_manual(values=c('#00BFC4', '#F8766D'))+
  theme_classic()+
  ylab(pi_exp)+
  ylim(0, 0.19)+
  geom_signif(stat='identity', 
                      data=data.frame(x = c(0.75, 1.75, 2.75, 3.75, 4.75),
                      xend=c(1.25, 2.25, 3.25, 4.25, 5.25),
                      y = c(0.175, 0.175, 0.175, 0.175, 0.175),
                      annotation2 = c('***',  ' *** ', '  ***  ', '   ***    ', '    ***    ')), 
              aes(x=x, xend=xend, y=y, yend=y, annotation=annotation2, textsize=2.5))+
    theme(legend.position='none', text=element_text(size=10), axis.title.x = element_blank())
  
p7 

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\pi_domain.png', plot=p7, dpi='retina', width=2.4, height=1.5)
```

Tajima's D: 
```{r}
cds <- compare_means(D~HV, col_popgen %>% filter(domain=='CDS'), method='wilcox.test', paired=FALSE)$p
cc <- compare_means(D~HV, col_popgen %>% filter(domain=='CC'), method='wilcox.test', paired=FALSE)$p
tir <- compare_means(D~HV, col_popgen %>% filter(domain=='TIR'), method='wilcox.test', paired=FALSE)$p
nbarc <- compare_means(D~HV, col_popgen %>% filter(domain=='NBARC'), method='wilcox.test', paired=FALSE)$p
lrr <- compare_means(D~HV, col_popgen %>% filter(domain=='LRR'), method='wilcox.test', paired=FALSE)$p

pvalues <- c(cds, cc, tir, nbarc, lrr)
p.adjust(pvalues, method = 'BH', n = length(pvalues))

compare_means(D~domain, col_popgen, method='wilcox.test')
compare_means(D~domain, col_popgen %>% filter(HV=='hv'), method='wilcox.test')
compare_means(D~domain, col_popgen %>% filter(HV=='non-hv'), method='wilcox.test')

kruskal.test(D ~ domain, data = col_popgen)
kruskal.test(D ~ domain, data = col_popgen%>%filter(HV=='hv'))
kruskal.test(D ~ domain, data = col_popgen%>%filter(HV=='non-hv'))
```


```{r}
p10 <- ggplot(col_popgen,
       aes(x=domain, y=D))+
   geom_boxplot(lwd=0.25, aes(fill=HV), outlier.size = 0.1)+
    scale_fill_manual(values=c('#00BFC4', '#F8766D'))+
  theme_classic()+
  ylab("Tajima's D")+
  ylim(-3, 4.4)+
  geom_signif(stat='identity', 
                      data=data.frame(x = c(0.75, 1.75, 2.75, 3.75, 4.75),
                      xend=c(1.25, 2.25, 3.25, 4.25, 5.25),
                      y = c(4.1, 4.1, 4.1, 4.1, 4.1),
                      annotation2 = c('***', '*', '**', ' *** ', '  ***  ')), 
              aes(x=x, xend=xend, y=y, yend=y, annotation=annotation2, textsize=2.5))+
    theme(legend.position='none', text=element_text(size=10), axis.title.x = element_blank())

  
p10 

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\D_domain.png', plot=p10, dpi='retina', width=2.2, height=1.5)

```

Pi vs Tajima's D plot 
```{r}
#calculate top 5% threshold for Pi and Tajima's D 
top_pi <- all_gene %>% pull(Pi_by_lseff) %>% quantile(c(.95), na.rm=TRUE)
top_D <- all_gene %>% pull(D) %>% quantile(c(.95), na.rm=TRUE)
all_gene %>% filter(HV != 'all_genes') %>% filter(Pi_by_lseff > top_pi & D > top_D)


p <-ggplot()+
  geom_density2d(all_gene, mapping=aes(x=Pi_by_lseff, y=D, linetype='Genome Distribution'), bins=20, color='grey',  lwd=0.5)+
  geom_segment(x=top_pi, 
               xend=top_pi, 
               y=max(all_gene$D, na.rm=TRUE), 
               yend=-3.1, color='black', 
               aes(linetype='95th Percentile'), lwd=0.35)+
  geom_segment(x=-1, 
               xend=max(all_gene$Pi_by_lseff, na.rm=TRUE), 
               y=top_D, 
               yend=top_D, color='black', 
               aes(linetype='95th Percentile'), lwd=0.35)+
  geom_point(data=to_plot, aes(x=Pi_site, y=D, color=HV), size=0.5)+
  scale_linetype_manual(values=c('Genome Distribution' = 1,  '95th Percentile'=2))+
  scale_color_manual(values = c("hv" = "#F8766D",
                                "non-hv" = '#00BFC4'))+
  guides(linetype=guide_legend(override.aes = list(linetype = c(2, 1),
                                                   shape=c(NA, NA), 
                                                   color=c('black', 'grey')), 
                               order=1),
         color=guide_legend(override.aes = list(linetype = c(0, 0), 
                                                shape=c(16, 16)), 
                            order=2), 
        fill = guide_legend(byrow = TRUE))+
  xlab(expression(pi))+
  ylab("Tajima's D")+
  theme_classic()+
  theme(legend.position='right', 
        legend.title=element_blank(), 
        legend.spacing.y=unit(-0.25, 'cm'),
        text=element_text(size=10))

p
ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\pi_v_d.png', plot=p, dpi='retina', width=5.75, height=2.75)
```



# Figure 4: 
PiN and PiS plots 
```{r}
pi_long <- to_plot %>% subset(select=c('Gene', 'PiS', 'PiN', 'HV')) %>% pivot_longer(
    cols = c(PiS, PiN), 
    names_to = "pi_type",
    values_to = "value") %>%
  mutate(pi_type=recode(pi_type, 
                        "PiS" = "{pi[S]}", 
                        "PiN" = "{pi[N]}"))

pis <- compare_means(value~HV, pi_long %>% filter(pi_type=="{pi[S]}"), method="wilcox.test")$p
pin <- compare_means(value~HV, pi_long %>% filter(pi_type=="{pi[N]}"), method="wilcox.test")$p

p.adjust(c(pis, pin), method='BH', 2)

pi_long$HV <- factor(pi_long$HV, levels=c('non-hv', 'hv'))

p13 <- pi_long %>%
  ggplot(aes(x=HV, y=value, fill=HV))+
  geom_violin(lwd=0.25)+
    geom_quasirandom(alpha=0.3, size=0.5)+
   geom_signif(comparisons=list(c('hv', 'non-hv')), 
               map_signif_level = TRUE, test=wilcox.test, textsize=2, size=0.25, y=c(0.08))+
  ylim(0,0.085)+
    scale_fill_manual(values=c('#00BFC4', '#F8766D'))+
  theme_classic()+
    theme(legend.position='none', text=element_text(size=10), axis.title=element_blank())+
  facet_grid(~pi_type, labeller=label_parsed)

p13

#ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\pisandpin.png', plot=p13, dpi='retina', width=2.5, height=2)
```

Supplemental figure 4: PiNPiS/domain
```{r}
cds <- compare_means(PiNPiS~HV, col_popgen %>% filter(domain=='CDS'), method='wilcox.test', paired=FALSE)$p
cc <- compare_means(PiNPiS~HV, col_popgen %>% filter(domain=='CC'), method='wilcox.test', paired=FALSE)$p
tir <- compare_means(PiNPiS~HV, col_popgen %>% filter(domain=='TIR'), method='wilcox.test', paired=FALSE)$p
nbarc <- compare_means(PiNPiS~HV, col_popgen %>% filter(domain=='NBARC'), method='wilcox.test', paired=FALSE)$p
lrr <- compare_means(PiNPiS~HV, col_popgen %>% filter(domain=='LRR'), method='wilcox.test', paired=FALSE)$p

pvalues <- c(cds, cc, tir, nbarc, lrr)
p.adjust(pvalues, method = 'BH', n = length(pvalues))

PiNPiS_domain <- ggplot(col_popgen,
       aes(x=domain, y=PiNPiS))+
   geom_boxplot(lwd=0.25, aes(fill=HV), outlier.size = 0.1)+
    scale_fill_manual(values=c('#00BFC4', '#F8766D'))+
  theme_classic()+
  ylab(expression(paste(pi[N],'/',pi[S])))+
  ylim(0, 3)+
  geom_signif(stat='identity', 
                      data=data.frame(x = c(0.75, 1.75, 2.75, 3.75, 4.75),
                      xend=c(1.25, 2.25, 3.25, 4.25, 5.25),
                      y = c(2.75, 2.75, 2.75, 2.75, 2.75),
                      annotation2 = c('n.s.', ' n.s. ', '  n.s.  ', '    n.s.   ', '    n.s.    ')), 
              aes(x=x, xend=xend, y=y, yend=y, annotation=annotation2, textsize=2.5))+
    theme(legend.position='none', text=element_text(size=10), axis.title.x = element_blank())

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\pinpis_domain.png', plot=PiNPiS_domain, dpi='retina', width=3, height=2.5)
```

PiN/PiS plot
```{r}
pis=expression(pi['S'])
pin=expression(pi['N'])
pis_equals_pin=expression(pi['s']~'='~pi['n'])



pis_piN <- to_plot %>% 
  ggplot(aes(x=PiS, y=PiN, fill=HV, color=HV))+
    geom_point(aes(color=HV), size=0.5)+
  geom_abline(slope=1, intercept=0, color='darkgrey', linetype='dashed')+
  geom_smooth(method='lm', se=FALSE, lwd=0.5)+
  ylim(0,0.08)+
  xlim(0,0.08)+
  scale_color_manual(values = c('#00BFC4',"#F8766D"))+
  xlab(pis)+
  ylab(pin)+
  theme_classic()+
  theme(text=element_text(size=10), 
        legend.text=element_text(size=10), 
        legend.title = element_blank(), 
        legend.key.size = unit(.2, 'cm'), 
        legend.key.height=unit(.2, 'cm'),
        legend.key.width = unit(.2, 'cm'))+
  annotate(geom="text", label=pis_equals_pin, x=0.06, y=0.0625, angle=40, vjust=-1, color='darkgrey', size=3.5)

pis_piN 

#ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\pis_pin.png', plot=pis_piN, dpi='retina', width=3.5, height=2.36)

#statistics and linear regression
to_plot %>% group_by(HV) %>% summarize(mean=mean(PiNPiS, na.rm=TRUE), median=median(PiNPiS, na.rm=TRUE))
compare_means(PiNPiS~HV, col_popgen%>%filter(domain=='CDS'), method='wilcox.test', paired=FALSE)
lm(PiN ~ PiS, data=to_plot %>% filter(HV=='hv'))
lm(PiN ~ PiS, data=to_plot %>% filter(HV=='non-hv'))

compare_means(PiNPiS~HV, col_popgen %>% filter(domain=='LRR'), method='wilcox.test', paired=FALSE)
col_popgen %>% filter(domain=='LRR') %>% group_by(HV) %>% summarize(mean=mean(PiNPiS, na.rm=TRUE), median=median(PiNPiS, na.rm=TRUE))


```


Mutation plot 
```{r}
mut_prob_plot <- NLR_only %>%
  ggplot(aes(x=HV, y=Mutation.Probability.Score, fill=HV))+
  geom_violin(lwd=0.25)+
  geom_quasirandom(alpha=0.3, size=0.5)+
  scale_fill_manual(values=c('#00BFC4', '#F8766D'))+
  theme_classic()+
  geom_signif(comparisons=list(c('non-hv', 'hv')), 
               map_signif_level = TRUE, test=wilcox.test, textsize=2, size=0.25)+
  theme(legend.position='none', text=element_text(size=10))+
  ylim(0,1.2e-4)+
  scale_y_continuous(labels = scales::scientific)+
  ylab("Mutation Probability")+
  xlab('')

mut_prob_plot 

#ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\mutprob.png', plot=mut_prob_plot, dpi='retina', width=3, height=2.5)
```


```{r}
compare_means(Mutation.Probability.Score~HV, NLR_only, method="wilcox.test")$p

summary <- NLR_only %>% group_by(HV) %>% summarize(mean=mean(Mutation.Probability.Score), median=median(Mutation.Probability.Score))
mean_percent_increase <- (summary[2,2]-summary[1,2])/summary[1,2]*100
med_percent_increase <- (summary[2,3]-summary[1,3])/summary[1,3]*100
print(mean_percent_increase[[1]])
print(med_percent_increase)
```

Positive selection sites plot: 
```{r}
facet_sites_internal <- sites_internal_long %>%
  ggplot(aes(x=HV, y=value*100, fill=HV))+
   geom_beeswarm(size=0.5, aes(color=HV))+
   geom_signif(comparisons=list(c('hv', 'non-hv')), 
               map_signif_level = TRUE, test=wilcox.test, textsize=2, size=0.25, y=c(16))+
    scale_color_manual(values=c('#00BFC4','#F8766D'))+
  theme_classic()+
  theme(legend.position='none', text=element_text(size=10))+
  facet_grid(~model, labeller=label_parsed)+
  ylab('% Codons Under + Selection')+
  xlab('')+
  ylim(0, 17)

p_pervasive <- compare_means(value~HV, sites_internal_long %>% filter(model=='Pervasive'), method='wilcox.test', paired=FALSE)$p
p_episodic <- compare_means(value~HV, sites_internal_long %>% filter(model=='Episodic'), method='wilcox.test', paired=FALSE)$p

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\sites_internal.png', plot=facet_sites_internal, dpi='retina', width=3.5, height=2.5)
```

```{r}
fig6 <- p13+pis_piN+mut_prob_plot+facet_sites_internal
ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\fig6.png', plot=fig6, dpi='retina', width=180, height=120, unit='mm')
```


# Figure 5: 
```{r}

col_popgen %>% filter(Gene=='AT5G43730')

#step 1: get AA coordinates of each domain from col_popgen, convert to nucleotide coordinates of coding sequence (multiply by 3)
annotation_creator <- function(gene_name){
  coords <- col_popgen %>% 
    filter(Gene==gene_name) %>% 
    filter(domain != 'CDS') %>%
    subset(select=c('Gene', 'domain', 'start', 'stop')) %>% 
    mutate(start=start*3)%>%
    mutate(stop=stop*3)
  
  return(coords)
}

gene_length_finder <- function(gene_name){
  gene_length <- col_popgen %>% 
    filter(Gene==gene_name) %>% 
    filter(domain == 'CDS') %>%
    pull(stop)
  
  return(gene_length*3)
}

coords <- annotation_creator('AT5G43730')
gene_length <- gene_length_finder('AT5G43730')

#step 2: plot
tripartite_plot <- function(gene_name){
  coords <- annotation_creator(gene_name)
  gene_length <- gene_length_finder(gene_name)
  
  #plot domains as rectangles
  domains <- ggplot()+
  theme_classic()+
  geom_rect(data=coords, aes(xmin=start, xmax=stop,
                             ymin=0, ymax=.2,  fill=domain),size=5)+
  scale_fill_manual(values=c('CC'='#00997F', 
                               'TIR'='orange',
                               'NBARC'='#00BADE', 
                               'LRR'='#9661FF'))+
  xlim(0, gene_length)+
  ylim(0,1)+
  theme(legend.position='right',
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.line.x=element_blank(), 
        axis.line.y=element_blank(), 
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  guides(fill=guide_legend(title="Domain"))
  
  D <- ggplot()+
  geom_point(data=col_slider %>% filter(Gene==gene_name) %>% filter(!is.na(D)), 
             aes(x=nuc_midpoint, y=D), size=1)+
  geom_line(data=col_slider %>% filter(Gene==gene_name) %>% filter(!is.na(D)), 
             aes(x=nuc_midpoint, y=D))+
  geom_vline(data=coords, linetype='dashed', aes(xintercept=start, color=domain))+
  geom_vline(data=coords, linetype='dashed', aes(xintercept=stop,  color=domain))+
   scale_color_manual(values=c('CC'='#00997F', 
                               'TIR'='orange',
                               'NBARC'='#00BADE', 
                               'LRR'='#9661FF'))+
  ylim(-3, 3)+
  xlim(0, gene_length)+
  ylab("D")+
  theme_classic()+
  theme(legend.position='none',
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        axis.line.x=element_blank())
  
  Pi <- col_slider %>% filter(Gene==gene_name) %>% ggplot() + 
    geom_point(aes(x=nuc_midpoint, y=Pi_site), size=1)+
    geom_line(aes(x=nuc_midpoint, y=Pi_site))+
  geom_vline(data=coords, linetype='dashed', aes(xintercept=start, color=domain))+
  geom_vline(data=coords, linetype='dashed', aes(xintercept=stop,  color=domain))+
  scale_color_manual(values=c('CC'='#00997F', 
                               'TIR'='orange',
                               'NBARC'='#00BADE', 
                               'LRR'='#9661FF'))+
  ylim(0, 0.1)+
  xlab('Nucleotide Position')+
  ylab(pi_exp)+
  xlim(0, gene_length)+
  theme_classic()+
  theme(legend.position='none')
  
  return(domains+D+Pi+plot_layout(guides='collect', ncol = 1))
 # return(Pi)
}
```


```{r}
AT5G43730 <- tripartite_plot('AT5G43730')
AT5G43740 <- tripartite_plot('AT5G43740')


ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\AT5G43730.png', plot=AT5G43730, dpi='retina', width=120, height=80, unit='mm')

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\AT5G43740.png', plot=AT5G43740, dpi='retina', width=120, height=80, unit='mm')
```

##Extended View 
RPP13 cluster: 'AT3G46530' 'AT3G46710' 'AT3G46730'
       Gene                 Clade HV
1 AT3G46730 Int14387_212_253_R_34  0
2 AT3G46710 Int14387_212_251_L_60  0
3 AT3G46530 Int14387_212_343_R_44  1

```{r}
#add interpro CC annotations since they are missing from pan-nlrome report...
m1 <- c('AT3G46730', NA, NA, NA, NA, NA, 'Int14387_212_253_R_34', NA, 'CC', NA, 3, 124, 'non-hv', NA, NA, NA, NA)
m2 <- c('AT3G46710', NA, NA, NA, NA, NA, 'Int14387_212_251_L_60', NA, 'CC', NA, 3, 124, 'non-hv', NA, NA, NA, NA)
m3 <- c('AT3G46530', NA, NA, NA, NA, NA, 'Int14387_212_343_R_44', NA, 'CC', NA, 3, 124, 'non-hv', NA, NA, NA, NA)
col_popgen <- rbind(col_popgen, m1) %>% rbind(m2) %>% rbind(m3)
col_popgen$start <- col_popgen$start %>% as.integer()
col_popgen$stop <- col_popgen$stop %>% as.integer()
```


```{r}
AT3G46530 <- tripartite_plot('AT3G46530')
AT3G46710 <- tripartite_plot('AT3G46710')
AT3G46730 <- tripartite_plot('AT3G46730')

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\AT3G46530.png', plot=AT3G46530, dpi='retina', width=4, height=3)

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\AT3G46710.png', plot=AT3G46710, dpi='retina', width=4, height=3)

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\AT3G46730.png', plot=AT3G46730, dpi='retina', width=4, height=3)
```



