---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)
library(openxlsx)
library(ComplexUpset)
library(ggsignif)
library(ggpubr)
```
ComplexUpset: potentially using to display the data in new and interesting ways 
```{r}
#read in the methylation, expression, and TE distance for all genes, filter to NLRs
all_gene_table <- readxl::read_xlsx(path="C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\all_gene_table.xlsx")
NLR_table <- all_gene_table %>% filter(HV != 'all_genes')

#read in the cluster information and N-term status 
cluster <- readxl::read_xlsx(path="C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\chae_2020_supplemental.xlsx")
table <- merge(NLR_table, cluster, by.x='Gene', by.y='gene', all.x=TRUE)
```

```{r}
clust <- table %>% filter(cluster_type != 'singleton') %>% filter(!is.na(meth_percentage))
compare_means(meth_percentage~HV, clust, method='t.test')
```

First, prepare the dataset. UpSet expects a boolean vectors under an indicator column header 

```{r}
table$HV <- table$HV == 'hv'
table$`non-HV` <- table$HV == FALSE
table$CNL <- table$Nterm == 'C'
table$TNL <- table$Nterm == 'T'
table$clustered <- table$cluster_type != 'singleton'
table$singleton <- table$cluster_type == 'singleton'
#variables: HV, CNL, TNL, RNL, clustered 

sets = c('HV', 'non-HV', 'clustered', 'CNL', 'TNL')
rev(sets)
```

Time to calculate the p value using a t test: 
```{r}
intersections=list('HV', 'non-HV', 
                         c('HV', 'clustered'), 
                         c('non-HV', 'clustered'),
                         c('HV', 'CNL'), 
                         c('non-HV', 'CNL'),
                         c('HV', 'TNL'), 
                         c('non-HV', 'TNL'), 
                         c('HV', 'clustered', 'CNL'), 
                         c('non-HV', 'clustered', 'CNL'),
                         c('HV', 'clustered', 'TNL'), 
                         c('non-HV', 'clustered', 'TNL')
                         )
comparisons <- c('hv_nhv', 'clust', 'cnl', 'tnl', 'tnl_clust', 'cnl_clust')

#transcription 
hv_nhv <- compare_means(log2_TPM~HV, table, method='t.test')$p.adj 
clust <- compare_means(log2_TPM~HV, table %>% filter(clustered == TRUE), method='t.test')$p.adj 
cnl <- compare_means(log2_TPM~HV, table %>% filter(Nterm == 'C'), method='t.test')$p.adj 
tnl <- compare_means(log2_TPM~HV, table %>% filter(Nterm == 'T'), method='t.test')$p.adj 
tnl_clust <- compare_means(log2_TPM~HV, table %>% filter(Nterm == 'T')%>%filter(clustered==TRUE), method='t.test')$p.adj
cnl_clust <- compare_means(log2_TPM~HV, table %>% filter(Nterm == 'C')%>%filter(clustered==TRUE), method='t.test')$p.adj

transcription_pval <- c(hv_nhv, clust, cnl, tnl, tnl_clust, cnl_clust)

#te_dist
a <- compare_means(te_dist~HV, table, method='t.test')$p.adj
b <- compare_means(te_dist~HV, table %>% filter(clustered == TRUE), method='t.test')$p.adj
c <- compare_means(te_dist~HV, table %>% filter(Nterm == 'C'), method='t.test')$p.adj
d <- compare_means(te_dist~HV, table %>% filter(Nterm == 'T'), method='t.test')$p.adj
e <- compare_means(te_dist~HV, table %>% filter(Nterm == 'T')%>%filter(clustered==TRUE), method='t.test')$p.adj
f <- compare_means(te_dist~HV, table %>% filter(Nterm == 'C')%>%filter(clustered==TRUE), method='t.test')$p.adj

te_dist_pval <- c(a, b, c, d, e, f)

#methylation
g <- compare_means(meth_percentage~HV, table, method='t.test')$p.adj
h <- compare_means(meth_percentage~HV, table %>% filter(clustered == TRUE), method='t.test')$p.adj
i <- compare_means(meth_percentage~HV, table %>% filter(Nterm == 'C'), method='t.test')$p.adj
j <- compare_means(meth_percentage~HV, table %>% filter(Nterm == 'T'), method='t.test')$p.adj
k <- compare_means(meth_percentage~HV, table %>% filter(Nterm == 'T')%>%filter(clustered==TRUE), method='t.test')$p.adj
l <- compare_means(meth_percentage~HV, table %>% filter(Nterm == 'C')%>%filter(clustered==TRUE), method='t.test')$p.adj

meth_pval <- c(g, h, i, j, k, l)

pval_table <- data.frame(comparisons, transcription_pval, meth_pval, te_dist_pval)
pval_table <- pval_table %>% mutate(transcription_stars = case_when(transcription_pval >= 0.05 ~ 'n.s.', 
                                                      transcription_pval < .05 & transcription_pval >= .01 ~ '*',
                                                      transcription_pval < .01 & transcription_pval >= .001 ~ '**',
                                                      transcription_pval < .001 & transcription_pval > 0 ~ '***')) %>%
  mutate(meth_stars = case_when(meth_pval >= 0.05 ~ 'n.s.', 
                                meth_pval < .05 & meth_pval >= .01 ~ '*',
                                meth_pval < .01 & meth_pval >= .001 ~ '**',
                                meth_pval < .001 & meth_pval > 0 ~ '***')) %>%
  mutate(te_stars = case_when(te_dist_pval >= 0.05 ~ 'n.s.', 
                                te_dist_pval < .05 & te_dist_pval >= .01 ~ '*',
                                te_dist_pval < .01 & te_dist_pval >= .001 ~ '**',
                                te_dist_pval < .001 & te_dist_pval > 0 ~ '***'))  
 
pval_table$transcription_pval <- pval_table$transcription_pval %>% as.character()
pval_table$meth_pval <- pval_table$meth_pval %>% as.character()
pval_table$te_dist_pval <- pval_table$te_dist_pval %>% as.character()
pval_table$transcription_stars

pval_table
```

Create the plot 
```{r}
p_upset2 <- upset(table, rev(sets), keep_empty_groups=TRUE, mode='inclusive_intersection', 
      sort_sets=FALSE,
      sort_intersections=FALSE, 
      intersections=list('HV', 'non-HV', 
                         c('HV', 'clustered'), 
                         c('non-HV', 'clustered'),
                         c('HV', 'CNL'), 
                         c('non-HV', 'CNL'),
                         c('HV', 'TNL'), 
                         c('non-HV', 'TNL'), 
                         c('HV', 'clustered', 'CNL'), 
                         c('non-HV', 'clustered', 'CNL'),
                         c('HV', 'clustered', 'TNL'), 
                         c('non-HV', 'clustered', 'TNL')
                         ),
      annotations=list('log2(TPM)' = (
        ggplot(mapping=aes(y=log2_TPM))+
                         geom_violin(aes(fill=intersection), lwd=0.25) +
          ylim(0,8) +
          geom_signif(stat='identity', 
                      data=data.frame(x = c(1, 3, 5, 7, 9, 11),
                      xend=c(2, 4, 6, 8, 10, 12),
                      y = c(7.5, 7.5, 7.5, 7.5, 7.5, 7.5),
                      annotation1 = pval_table$transcription_stars, 
                      annotation2 = c('***', ' *** ', '  ***  ', '**', '*', '   ***   ')), 
              aes(x=x, xend=xend, y=y, yend=y, annotation=annotation2, textsize=3))+
          theme(text=element_text(size=10), legend.position='none')
        ),
  '%CG Methylation' = (
        ggplot(mapping=aes(y=meth_percentage))+
          geom_violin(aes(fill=intersection), lwd=0.25)+
        #  theme_classic()+
        theme(text=element_text(size=10), legend.position = 'none')+
          ylim(0, 100)+
        geom_signif(stat="identity", 
              data=data.frame(x = c(1, 3, 5, 7, 9, 11),
                              xend=c(2, 4, 6, 8, 10, 12),
                              y = c(90, 90, 90, 90, 90, 90),
                              annotation = c('**', 'n.s.', '  *  ', '  n.s.  ', ' n.s. ', ' * ')), 
              aes(x=x, xend=xend, y=y, yend=y, annotation=annotation, textsize=3))
        )
    , 
      'TE distance (kb)' = (
        ggplot(mapping=aes(y=te_dist))+
          geom_boxplot(aes(fill=intersection), lwd=0.25)+
         # theme_classic()+
          theme(text=element_text(size=10), legend.position = 'none')+
          ylim(0, 38)+
          geom_signif(stat="identity", 
              data=data.frame(x = c(1, 3, 5, 7, 9, 11),
                              xend=c(2, 4, 6, 8, 10, 12),
                              y = c(34, 34, 34, 34, 34, 34),
                              annotation = c('***', ' *** ', '  ***  ', '**', '   ***   ', ' ** ')), 
              aes(x=x, xend=xend, y=y, yend=y, annotation=annotation, textsize=3), tip_length=0.1))
        ),   
  themes=upset_default_themes(text=element_text(size=10)) , 
   base_annotations=list('Inclusive intersection size'=(
     intersection_size(mode='inclusive_intersection', text=list(size=3)) + ylab('Set Size')))
  )

p_upset2
ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\e14_R\\final_figures\\upset_sig.png', plot=p_upset2, dpi='retina', width=5, height=6.5)
```

Solution: looks like the annotations have to be unique for some reason? so lots of balanced spaces 

What I want to do with this summary figure, however, is do sets of "highly expressed", "not expressed" or "methylated" and "not methylated", "contains TE", does not contain TE and then the proportion? of hv to nonhv (but it has to be relative to small HV set size ) 