---
title: "Violin plots"
author: "Chandler Sutherland"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(ggplot2)
library(ggsignif)
library(ggpubr)
library(ggbeeswarm)
library(patchwork)
```

Copyright (c) Chandler Sutherland
Email: chandlersutherland@berkeley.edu

Purpose: generate hv vs non-hvNLR violin plots of expression vs methylation.

Intermediate processing steps are shown here, but figures can be recreated using just the numerical source data provided in Source `Data/Figure 2`.

```{r}
#input data 
table <- readxl::read_xlsx(path="C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E14\\all_gene_table.xlsx")

NLR_table <- table %>% filter(HV != 'all_genes')
NLR_table$HV <- factor(NLR_table$HV , levels=c("non-hv", "hv"))
fig2_source <- NLR_table %>% subset(select=c('Gene', 'HV', 'log2_TPM', 'meth_percentage', 'te_dist'))
```


```{r}
#save source data. To reproduce these plots, un-comment reading in the fig2_source and start code from there  
write.csv(fig2_source, file="./Source Data/Figure 2/2A-C/per_NLR_features.csv")
#fig2_source <- read.csv("./Source Data/Figure 2/2A-C/per_NLR_features.csv")
```

Fig 2a: Expression in log2(TPM)
```{r}
#create a sample size label
NLR_table2 <- fig2_source %>% 
  filter(!is.na(log2_TPM)) %>% 
  group_by(HV) %>% 
  mutate(n_exp = n()) %>% 
  mutate(exp_label = paste(HV, n_exp, sep="\nn="))
levels <- NLR_table2 %>% pull(exp_label) %>% unique()
NLR_table2$exp_label <- factor(NLR_table2$exp_label, levels=c(levels[1], levels[2]))

#plot 
p1 <- ggplot(NLR_table2,
       aes(x=exp_label, y=log2_TPM, fill=HV))+
    geom_violin(lwd=0.30)+
  geom_quasirandom(alpha=0.3, size=0.5)+
   ylim(0,8.8)+
   geom_signif(comparisons=list(c(levels[1], levels[2])), 
               map_signif_level = TRUE, y_position = c(8, 7), test=wilcox.test, textsize=2, size=0.25)+
    scale_fill_manual(values=c('#00BFC4',  '#F8766D'))+
  theme_classic()+
    theme(legend.position='none', text=element_text(size=10), 
          axis.title=element_blank())+
  scale_x_discrete(labels=c(levels[1], levels[2]))

p1

#unpaired wilcox rank sum test 
compare_means(log2_TPM~HV, NLR_table, method = 'wilcox.test', paired = FALSE)
```

Fig 2B: %CG methylation 
```{r}
# sample size label
NLR_table3 <- fig2_source %>% filter(!is.na(meth_percentage)) %>% group_by(HV) %>% mutate(n_meth = n()) %>% mutate(meth_label = paste0(HV,"\n", "n=",n_meth))
levels_m <- NLR_table3 %>% pull(meth_label) %>% unique()
NLR_table3$meth_label <- factor(NLR_table3$meth_label, levels=c(levels_m[1], levels_m[2]))

#plot
p2 <- ggplot(NLR_table3,
       aes(x=meth_label, y=meth_percentage, fill=HV))+
    geom_violin(lwd=0.3)+
  geom_quasirandom(alpha=0.3, size=0.5)+
  ylim(0,100)+
   geom_signif(comparisons=list(c(levels_m[1], levels_m[2])), 
               map_signif_level = TRUE, y_position = c(90, 80), test=wilcox.test, textsize=2, size=0.25)+
    scale_fill_manual(values=c('#00BFC4',  '#F8766D'))+
  theme_classic()+
    theme(legend.position='none', text=element_text(size=10), axis.title=element_blank())

p2

#stats 
compare_means(meth_percentage~HV, NLR_table, method = 'wilcox.test', paired = FALSE)
```

Fig 2C: Distance to nearest TE 
```{r}
#sample size label
NLR_table3 <- fig2_source %>% 
  filter(!is.na(te_dist)) %>% 
  group_by(HV) %>% 
  mutate(n_te = n()) %>% 
  mutate(te_label = paste0(HV,"\n", "n=",n_te))
levels_t <- NLR_table3 %>% pull(te_label) %>% unique()
NLR_table3$te_label <- factor(NLR_table3$te_label, levels=c(levels_t[1], levels_t[2]))

#plot
p3 <- ggplot(NLR_table3,
       aes(x=te_label, y=te_dist, fill=HV))+
    geom_violin(lwd=0.3)+
   ylim(0,35)+
  geom_quasirandom(alpha=0.3, size=0.5)+
   geom_signif(comparisons=list(c('non-hv\nn=97', 'hv\nn=35')), 
               map_signif_level = TRUE, y_position = c(32, 32), test=wilcox.test, textsize=2, size=0.25)+
    scale_fill_manual(values=c('#00BFC4',  '#F8766D'))+
  theme_classic()+
    theme(legend.position='none', text=element_text(size=10), 
          axis.title=element_blank())

#unpaired wilcox rank sum test 
compare_means(te_dist~HV, NLR_table, method = 'wilcox.test', paired = FALSE)
```

```{r}
#perform fisher's test on HV status vs internal TE
NLR_table3 <- NLR_table3 %>% mutate('int_te'=if_else(te_dist == 0, 'TRUE', 'FALSE'))
fisher.test(NLR_table3$HV, NLR_table3$int_te)
```
```{r}
#combine plots and save 
panel <- p1+p2+p3

ggsave(filename='C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\Outputs\\NLR Features Paper\\EMBO Submission\\Figure Panels\\fig_2a_c.svg', plot=panel, dpi=1000, width=180, height=50, unit='mm')
```

